\section{Applicazioni BIM nel Web}
\label{sec:chapter_1_section_3}

Nowadays we are seeing a relentless migration of software products toward services accessible via the Web medium. This is mainly due to the undeniable benefits in terms of accessibility, usability, maintainability and spreadability granted by the Web medium itself. Nevertheless these benefits don't come without a cost: performance  and development complexity become major concerns in the Web environment.

In particular, due to the introduction of several abstraction layers it is not always feasible to "port" a desktop application into the Web realm, an aspect to be taken into account even for the relevant hardware differences among all the devices equipped with a Web Browser. It can be even more arduous to tackle the inherent distributed software architecture (a client/server one at least) induced by the Web platform. Nevertheless increasingly rich and complex Web applications began to appear, supported by the enriched HTML5 APIs, which thanks to the WebGL~\cite{webgl} (which enables direct access to GPU), Canvas~\cite{Munro:15:HCC} (2D raster APIs) and SVG~\cite{Jackson:11:SVG} (vectorial drawing APIs), has paved the way for the entrance of Web Graphic Applications.

In this work we report about our endeavor toward the definition of a Web based buildings modeling tool which overcomes the aforementioned performance and development difficulties relying on a unidirectional data flow design pattern and on a serverless architecture, respectively.

A serverless architecture, on the contrary of what the name may suggest, actually employs many different specific servers, whose operation and maintenance don't burden on the project developer(s). These several servers can be seen as third party services (typically cloud-based) or functions executed into ephemeral containers (may only last for one invocation) to manage the internal state and server-side logic. Realtime interaction among users jointly working on the same modeling project, is for example achieved via a third party APIs for remote users collaboration.

The tool user interface, entirely based on web components pattern,  has been kept as simple as possible: the user is required to interact mainly with two-dimensional symbolic placeholders representing parts of the building, thus avoiding complex 3D interactions.
The modeling complexity is thus moved from the modeler to the developer which fills out an extendible \emph{catalog} of customizable \emph{building elements}. The modeler has only to select the required element, place and parametrize it according to the requirements. It is obvious that a large number of building elements has to be provided to ensure the fulfillment of the most modeling requirements.

The remainder of this document is organized as follows. Section~\ref{sec:related_work} provides an overview of related work. Section~\ref{sec:application} reports about the application user experience. Section~\ref{sec:architecture} presents adopted architectural solutions. Finally, Section ~\ref{sec:conclusions} contains some conclusive remarks.
